---
title: "Graphs and Box Plots of The Optimal Genes"
author: "Basak"
date: "7/2/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
library(dplyr)
library(tibble)
library(caret)
library(ggplot2)
library(mlbench)
require(caret)
require(randomForest)
require(Boruta)
library(pROC)
require(ppcor)
library(qgraph)
library(ggpubr)
```

### Reading ranked gene sets 
```{r}
# your data
rank1 = read.csv('top_list1-8.csv', header = TRUE)
rank2 = read.csv('top_list2-8.csv', header = TRUE)
rank3 = read.csv('top_list3-8.csv', header = TRUE)

genes = rbind(rank1,rank2,rank3)
genes = unique(genes$x)
```
### Reading data sets
###### HCCDB3

```{r}
x.3 = read.csv('DEGS_HCCDB3.csv', row.names = 1, header = TRUE)
x.3 = t(x.3)
x.3 = as.data.frame(scale(x.3,center = TRUE, scale = TRUE))
x.3[1:5,1:5]

sample3_info = read.csv('prepared_sample_info3.csv', header = TRUE) 

head(sample3_info)

```
###### HCCDB4

```{r}
x.4 = read.csv('DEGS_HCCDB4.csv', row.names = 1, header = TRUE)
x.4 = t(x.4)
x.4 = as.data.frame(scale(x.4,center = TRUE, scale = TRUE))
x.4[1:5,1:5]

sample4_info = read.csv('prepared_sample_info4.csv', header = TRUE) 
# head(sample4_info)

```
###### HCCDB6

```{r}
x.6 = read.csv('DEGS_HCCDB6.csv', row.names = 1, header = TRUE)
x.6 = t(x.6)
x.6 = as.data.frame(scale(x.6,center = TRUE, scale = TRUE))
x.6[1:5,1:5]

sample6_info = read.csv('prepared_sample_info6.csv', header = TRUE) 
# head(sample6_info)

```

### Preparing data sets
```{r}
data3 = x.3 %>% 
  rownames_to_column(var = "Samples") %>%
  left_join(sample3_info %>% dplyr::select(Samples, TYPE), by = "Samples") %>%
  column_to_rownames(var = "Samples") %>%
  dplyr::select(TYPE, genes)
# data3[1:5,1:5]
```

```{r}
data4 = as.data.frame(x.4) %>% 
  rownames_to_column(var = "Samples") %>%
  left_join(sample4_info %>% dplyr::select(Samples,TYPE), by = "Samples") %>%
  column_to_rownames(var = "Samples") %>%
  dplyr::select(TYPE,genes)
# data4[1:5,1:5]
```

```{r}
data6 = as.data.frame(x.6) %>% 
  rownames_to_column(var = "Samples") %>%
  left_join(sample6_info %>% dplyr::select(Samples,TYPE), by = "Samples") %>%
  column_to_rownames(var = "Samples") %>%
  dplyr::select(TYPE,genes)
# data6[1:5,1:5]
```

## HCCDB3
#### Adjacent sample correlations
###### By using cor_auto with pcor in qgraph package
```{r}
adj.data3 = data3 %>% filter(TYPE=='Adjacent') %>% dplyr::select(-TYPE)
adj.pcor3 = cor_auto(adj.data3)
pdf(file = "Partial Correlation Graph - HCCDB3 - Adjacent.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5) 
graph_adj.pcor3 <- qgraph(adj.pcor3, graph = "pcor", layout = "spring", threshold = "BH",
                     sampleSize = nrow(adj.data3), alpha = 0.05, 
                     labels = colnames(adj.pcor3),
                     label.cex = 1.2,
                     vsize = 8,
                     label.font = 2
                     # title = 'HCCDB3 - Adjacent Sample Gene Network'
                     )
dev.off()

```



#### HCC sample correlations
###### By using cor_auto with pcor in qgraph package
```{r}
hcc.data3 = data3 %>% filter(TYPE=='HCC') %>% dplyr::select(-TYPE)
hcc.pcor3 = cor_auto(hcc.data3)
pdf(file = "Partial Correlation Graph - HCCDB3 - HCC.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5) 
graph_hcc.pcor3 <- qgraph(hcc.pcor3, graph = "pcor", layout = "spring", threshold = "BH",
                     sampleSize = nrow(hcc.data3), alpha = 0.05, 
                     labels = colnames(hcc.pcor3),
                     label.cex = 1.2,
                     vsize = 8,
                     label.font = 2
                     # title = 'HCCDB3 - HCC Sample Gene Network'
                     )
dev.off()
```

```{r}
######## This Code chunk to obtain the number of connections that each gene has.########
#  Number of edges of genes (Adjacent)
# Genes are stored as Numbers but qgraph converts them internally by input from user --> colnames(adj.pcor3)
# This code uses this knowledge to convert numbes to node names since Edgelist varible is stores as numbers 
# before converted into gene names.
# This code gets number of edges for each gene
connections.adj = as.data.frame(sort(table(c(graph_adj.pcor3$Edgelist$from, graph_adj.pcor3$Edgelist$to)), decreasing = TRUE))
Var1 = 1:length(colnames(adj.pcor3))
names.adj = as.data.frame(Var1) %>% mutate(Genes.adj = c(colnames(adj.pcor3))) %>% mutate(Var1 = as.factor(Var1))
connections.adj = connections.adj %>% left_join(names.adj, by = 'Var1')
connections.adj

#  Number of edges of genes (HCC)
connections.hcc = as.data.frame(sort(table(c(graph_hcc.pcor3$Edgelist$from, graph_hcc.pcor3$Edgelist$to)), decreasing = TRUE))
Var1 = 1:length(colnames(hcc.pcor3))
names.hcc = as.data.frame(Var1) %>% mutate(Genes.hcc = c(colnames(hcc.pcor3))) %>% mutate(Var1 = as.factor(Var1))
connections.hcc = connections.hcc %>% left_join(names.hcc, by = 'Var1')
connections.hcc

connections = connections.adj %>% left_join(connections.hcc, by = c('Var1')) 
# As it can be seen Genes.hcc and Genes.adj are same and have same Var1 name (sanity check)
connections = connections %>% mutate(difference = Freq.x - Freq.y)
connections %>% arrange(desc(difference))
```



#### Box-Plots of Genes
```{r}
# Specify the comparisons you want
my_comparisons <- list( c("Adjacent", "HCC") )

 for(i in 2:23) { 
  nam <- paste("g", i, sep = "")
  y = colnames(data3)[i]
  assign(nam, ggboxplot(data3,x = 'TYPE', y = y,color = 'TYPE',
                        palette =c("#FC4E07","#00AFBB","#FC4E07"),shape = 'TYPE') + 
                        stat_compare_means(label.x = 1.2, label.y = NULL) )
 }
```

###### Down Regulated
```{r}
pdf(file = "Down Regulated-HCCDB3_1.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
# For data 3
ggarrange(g3, g4, g5, g6, g7 , g9, heights = c(20, 20),
             ncol = 3, nrow = 2, labels = c("A","B", "C","D","E","F"))
dev.off()

pdf(file = "Down Regulated-HCCDB3_2.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange( g10, g11, g13, g14, g15, g16, heights = c(10, 10),
             ncol = 3, nrow = 2, labels = c("A","B", "C","D","E","F"))
dev.off()

pdf(file = "Down Regulated-HCCDB3_3.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange(g17,g18,g19, g21, g22,g12, heights = c(10, 10),
             ncol = 3, nrow = 2, labels = c("A","B", "C","D","E","F"))
dev.off()
```


###### Up Regulated
```{r}
pdf(file = "Up Regulated-HCCDB3.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange(g2, g8, g20, g23, heights = c(10, 10),
          ncol = 2, nrow = 2, labels = c("A","B", "C","D"))
dev.off()
```

## HCCDB4
#### Adjacent sample correlations
###### By using cor_auto with pcor in qgraph package
```{r}
adj.data4 = data4 %>% filter(TYPE=='Adjacent') %>% dplyr::select(-TYPE)
adj.pcor4 = cor_auto(adj.data4)
pdf(file = "Partial Correlation Graph - HCCDB4 - Adjacent.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5) 
graph_adj.pcor4 <- qgraph(adj.pcor4, graph = "pcor", layout = "spring", threshold = "BH",
                     sampleSize = nrow(adj.data4), alpha = 0.05, 
                     labels = colnames(adj.pcor4),
                     label.cex = 1.2,
                     vsize = 8,
                     label.font = 2
                     # title = 'HCCDB4 - Adjacent Sample Gene Network'
                     )
dev.off()
```



#### HCC sample correlations
###### By using cor_auto with pcor in qgraph package
```{r}
hcc.data4 = data4 %>% filter(TYPE=='HCC') %>% dplyr::select(-TYPE)
hcc.pcor4 = cor_auto(hcc.data4)
pdf(file = "Partial Correlation Graph - HCCDB4 - HCC.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5) 
graph_hcc.pcor4 <- qgraph(hcc.pcor4, graph = "pcor", layout = "spring", threshold = "BH",
                     sampleSize = nrow(hcc.data4), alpha = 0.05, 
                     labels = colnames(hcc.pcor4),
                     label.cex = 1.2,
                     vsize = 8,
                     label.font = 2
                     # title = 'HCCDB4 - HCC Sample Gene Network'
                     )
dev.off()
```
```{r}

# Number of edges of genes (Adjacent)
# Genes are stored as Numbers but qgraph converts them internally by input from user --> colnames(adj.pcor3)
# This code uses this knowledge to convert numbes to node names since Edgelist varible is stores as numbers before converted into gene names.
# This code gets number of edges for each gene
connections.adj = as.data.frame(sort(table(c(graph_adj.pcor4$Edgelist$from, graph_adj.pcor4$Edgelist$to)), decreasing = TRUE))
Var1 = 1:length(colnames(adj.pcor4))
names.adj = as.data.frame(Var1) %>% mutate(Genes.adj = c(colnames(adj.pcor4))) %>% mutate(Var1 = as.factor(Var1))
connections.adj = connections.adj %>% left_join(names.adj, by = 'Var1')
connections.adj

#  Number of edges of genes (HCC)
connections.hcc = as.data.frame(sort(table(c(graph_hcc.pcor4$Edgelist$from, graph_hcc.pcor4$Edgelist$to)), decreasing = TRUE))
Var1 = 1:length(colnames(hcc.pcor4))
names.hcc = as.data.frame(Var1) %>% mutate(Genes.hcc = c(colnames(hcc.pcor4))) %>% mutate(Var1 = as.factor(Var1))
connections.hcc = connections.hcc %>% left_join(names.hcc, by = 'Var1')
connections.hcc

connections = connections.adj %>% left_join(connections.hcc, by = c('Var1')) 
# As it can be seen Genes.hcc and Genes.adj are same and have same Var1 name (sanity check)
connections = connections %>% mutate(difference = Freq.x - Freq.y)
connections %>% arrange(desc(difference))
```


#### Box-Plots of Genes
```{r}
# Specify the comparisons you want
my_comparisons <- list( c("Adjacent", "HCC") )

 for(i in 2:23) { 
  nam <- paste("g", i, sep = "")
  y = colnames(data4)[i]
  assign(nam, ggboxplot(data4,x = 'TYPE', y = y,color = 'TYPE',
                        palette =c("#FC4E07","#00AFBB","#FC4E07"),shape = 'TYPE') + 
                        stat_compare_means(label.x = 1.2, label.y = NULL) )
 }
```

###### Down Regulated
```{r}

# For data 4
pdf(file = "Down Regulated-HCCDB4_1.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange(g3, g4, g5, g6, g7 , g9, heights = c(20, 20),
             ncol = 3, nrow = 2, labels = c("A","B", "C","D","E","F"))
dev.off()

pdf(file = "Down Regulated-HCCDB4_2.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange( g10, g11, g13, g14, g15, g16, heights = c(10, 10),
             ncol = 3, nrow = 2, labels = c("A","B", "C","D","E","F"))
dev.off()

pdf(file = "Down Regulated-HCCDB4_3.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange(g17,g18,g19, g21, g22,g12, heights = c(10, 10),
             ncol = 3, nrow = 2, labels = c("A","B", "C","D","E","F"))
dev.off()
```


###### Up Regulated
```{r}
pdf(file = "Up Regulated-HCCDB4.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange(g2, g8, g20, g23, heights = c(10, 10),
          ncol = 2, nrow = 2, labels = c("A","B", "C","D"))
dev.off()
```



## HCCDB6
#### Adjacent sample correlations
###### By using cor_auto with pcor in qgraph package
```{r}
adj.data6 = data6 %>% filter(TYPE=='Adjacent') %>% dplyr::select(-TYPE)
adj.pcor6 = cor_auto(adj.data6)
pdf(file = "Partial Correlation Graph - HCCDB6 - Adjacent.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5) 
graph_adj.pcor6 <- qgraph(adj.pcor6, graph = "pcor", layout = "spring", threshold = "BH",
                     sampleSize = nrow(adj.data6), alpha = 0.05, 
                     labels = colnames(adj.pcor6),
                     label.cex = 1.2,
                     vsize = 8,
                     label.font = 2
                     # title = 'HCCDB6 - Adjacent Sample Gene Network'
                     )
dev.off()
```


#### HCC sample correlations
###### By using cor_auto with pcor in qgraph package
```{r}
hcc.data6 = data6 %>% filter(TYPE=='HCC') %>% dplyr::select(-TYPE)
hcc.pcor6 = cor_auto(hcc.data6)
pdf(file = "Partial Correlation Graph - HCCDB6 - HCC.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5) 
graph_hcc.pcor6 <- qgraph(hcc.pcor6, graph = "pcor", layout = "spring", threshold = "BH",
                     sampleSize = nrow(hcc.data6), alpha = 0.05, 
                     labels = colnames(hcc.pcor6),
                     label.cex = 1.2,
                     vsize = 8,
                     label.font = 2
                     # title = 'HCCDB6 - HCC Sample Gene Network'
                     )
dev.off()
```
```{r}

#  Number of edges of genes (Adjacent)
# Genes are stored as Numbers but qgraph converts them internally by input from user --> colnames(adj.pcor3)
# This code uses this knowledge to convert numbes to node names since Edgelist varible is stores as numbers before converted into gene names.
# This code gets number of edges for each gene
connections.adj = as.data.frame(sort(table(c(graph_adj.pcor6$Edgelist$from, graph_adj.pcor6$Edgelist$to)), decreasing = TRUE))
Var1 = 1:length(colnames(adj.pcor6))
names.adj = as.data.frame(Var1) %>% mutate(Genes.adj = c(colnames(adj.pcor6))) %>% mutate(Var1 = as.factor(Var1))
connections.adj = connections.adj %>% left_join(names.adj, by = 'Var1')
connections.adj

#  Number of edges of genes (HCC)
connections.hcc = as.data.frame(sort(table(c(graph_hcc.pcor6$Edgelist$from, graph_hcc.pcor6$Edgelist$to)), decreasing = TRUE))
Var1 = 1:length(colnames(hcc.pcor6))
names.hcc = as.data.frame(Var1) %>% mutate(Genes.hcc = c(colnames(hcc.pcor6))) %>% mutate(Var1 = as.factor(Var1))
connections.hcc = connections.hcc %>% left_join(names.hcc, by = 'Var1')
connections.hcc

connections = connections.adj %>% left_join(connections.hcc, by = c('Var1')) 
# As it can be seen Genes.hcc and Genes.adj are same and have same Var1 name (sanity check)
connections = connections %>% mutate(difference = Freq.x - Freq.y)
connections %>% arrange(desc(difference))
```

#### Box-Plots of Genes
```{r}
# Specify the comparisons you want
my_comparisons <- list( c("Adjacent", "HCC") )

 for(i in 2:23) { 
  nam <- paste("g", i, sep = "")
  y = colnames(data6)[i]
  assign(nam, ggboxplot(data6, x = 'TYPE', y = y, color = 'TYPE',
                        palette =c("#FC4E07","#00AFBB","#FC4E07"),shape = 'TYPE') + 
                        stat_compare_means(label.x = 1.2, label.y = NULL) )
 }
```

###### Down Regulated
```{r}

# For data 6
pdf(file = "Down Regulated-HCCDB6_1.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange(g3, g4, g5, g6, g7 , g9, heights = c(20, 20),
             ncol = 3, nrow = 2, labels = c("A","B", "C","D","E","F"))
dev.off()

pdf(file = "Down Regulated-HCCDB6_2.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange( g10, g11, g13, g14, g15, g16, heights = c(10, 10),
             ncol = 3, nrow = 2, labels = c("A","B", "C","D","E","F"))
dev.off()

pdf(file = "Down Regulated-HCCDB6_3.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange(g17,g18,g19, g21, g22,g12, heights = c(10, 10),
             ncol = 3, nrow = 2, labels = c("A","B", "C","D","E","F"))
dev.off()
```


###### Up Regulated
```{r}
pdf(file = "Up Regulated-HCCDB6.pdf",
    width = 11, # The width of the plot in inches
    height = 8.5)
ggarrange(g2, g8, g20, g23, heights = c(10, 10),
          ncol = 2, nrow = 2, labels = c("A","B", "C","D"))
dev.off()
```


