---
title: "HCCDB4"
author: "Basak"
date: "6/25/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(MASS)
library("AnnotationDbi")
library("org.Hs.eg.db")
library(dplyr)
library(tibble)
library("factoextra")
```

## DATA PREPARATION
### Reading the related HCCDB3 data sets 
```{r}
setwd('/Users/user/Documents/MSc BIOINFORMATICS/Individual Project/HCCDB4')
sample_info = read.table('HCCDB4.sample.txt', header = TRUE, row.names = 1)
sample_info = as.data.frame(t(sample_info))
sample_info = sample_info %>% rownames_to_column(var = 'Samples')
head(sample_info)
table(sample_info$TYPE)
```


```{r}
data = read.csv('GSE36376.gene.txt',sep = '\t',header = TRUE)
data[1:5,1:5]
```


### Checking Gene Names and ENTREZ ID Frequencies
Below code chunk checks frequencies of the genes in "Symbol" column of the data. Moreover, the second chunck checks the uniqueness of ENTREZ_ID column, as well, since it will be used to re-annotate the symbols later. 

```{r}
# Occurence of genes in the dataset
counts = count(data, data$Symbol)
# Genes which occur more than one
counts[counts$n>1,]
# NA which occur more than one in the Symbol column
counts[is.na(counts$`data$Symbol`),]
```

```{r}
# Occurence of ENTREZ ID's in the dataset
counts = count(data, data$Entrez_ID)
# ENTREZ ID's which occur more than one
counts[counts$n>1,]
# NA which occur more than one in the ENTREZ ID column
counts[is.na(counts$`data$Entrez_ID`),]
```

### Gene symbols are re-annotated by using their ENTREZ ID. This is also needed because later when these symbols are used in immune deconvolution algorithms, some methods require specifically "AnnotationDbi" annotated gene symbols.
```{r}
geneSymbols =  AnnotationDbi::select(org.Hs.eg.db, keys=as.character(data$Entrez_ID),
                                     columns = c("ENTREZID", "SYMBOL"),keytype="ENTREZID",
                                     multiVals="first")
# Match and replace symbols with re-annotated symbols in the data
data$Symbol[match(geneSymbols$ENTREZID, data$Entrez_ID)] <- geneSymbols$SYMBOL
```
### Checking Gene Names Frequency again, and removing the ones which can not be annotated.
```{r}
# Occurence of genes in the dataset
counts = count(data, data$Symbol)
# Gene names which occur more than one 
counts[counts$n>1,]
# NA which occur more than one in the Symbol column
counts[is.na(counts$`data$Symbol`),]
```

```{r}
# remove rows with no gene names (removes NAs)
data = data[complete.cases(data$Symbol),]
rownames(data) = data$Symbol
```

### Matching the order of sample info and data (Checking)
```{r}
# match the sample names
match_index = match(sample_info$Samples, colnames(data))
data = data[,match_index]
dim(data)
data[1:5,1:5]
```
### Checking Missing Values
```{r}
which(is.na(data), arr.ind=TRUE)
```
### Data is already log2 transformed. 
```{r}
boxplot(data[,runif(10, 1, dim(data)[2])])
```
### Transpose data for missing imputation with its related group mean, but also for PCA since it requires rows of the data to be observations (n=patients) and columns to be variables(p=genes). The coefficient matrix is p by p matrix
```{r}
data = as.data.frame(t(data))
data[1:5,1:5]
```
### Imputing missing values. 
#### Note: Although, output shows some messages about 'ignoring the group variable column', it does the imputation by considering grouping, but skips that grouping column when performing missing value imputation.
```{r}
data = data %>% 
  rownames_to_column(var = "Samples") %>%
  left_join(sample_info %>% select(Samples,TYPE), by = "Samples") %>%
  group_by(TYPE) %>%
  mutate_all(funs(ifelse(is.na(.), mean(., na.rm = TRUE),.))) %>%
  column_to_rownames(var = "Samples")
# Remove TYPE variable for further analysis
data = data %>% select(-TYPE)
```

### Data is already log2 transformed. Scaling is applied in order to fix the symmetry of the data and make it more like normally distributed in order to perform PCA.

```{r}
groups = sample_info$TYPE
res.pca <- prcomp(data, scale. = TRUE) # with autoscaling
#library(ggplot2)
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = groups, # color by groups
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups", label = "ind"
             ) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
r=res.pca$x

pdf(file = "HCCDB4 - PCA.pdf")
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = groups, # color by groups
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups", label = "ind"
             ) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
dev.off()
```

### By visual inspection, it can be seen that one Adjacent sample sits inside of HCC cluster. It has been removed with the following codes, since it might be potentially 

```{r}
outlier = which(r[1:193,1]>0 & r[1:193,2]>0)
res.pca <- prcomp(data[-outlier,], scale. = TRUE)
#library(ggplot2)
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = groups[-outlier], # color by groups
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups", label = "ind"
             ) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
pdf(file = "HCCDB4 - PCA- Without Mislabeled One.pdf")
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = groups[-outlier], # color by groups
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups", label = "ind"
             ) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
dev.off()

```

### Write out (un-scaled) HCCDB4 expression data set and sample data set into csv file.
#### Note: Unscaled is needed for immune deconvolution algorithms. One algoritm requires even without log transformation which will be undone later on.
```{r}
write.csv(data[-outlier,], 'prepared_data4.csv',quote = F,row.names = T)
write.csv(sample_info[-outlier,], 'prepared_sample_info4.csv',quote = F,row.names = F)
```

