---
title: "Random Forest with Selected Features Combination 1"
author: "Basak"
date: "6/29/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
library(dplyr)
library(tibble)
library(caret)
library(ggplot2)
library(mlbench)
require(caret)
require(randomForest)
require(Boruta)
library(pROC)
```
### Reading ranked genes data sets 
```{r}
rank3 = read.csv('new_boruta_ranking_data3.csv', row.names = 1, header = TRUE)
rank4 = read.csv('new_boruta_ranking_data4.csv', row.names = 1, header = TRUE)


# filtering confirmed Boruta selected features
rank3 = rank3 %>% 
  rownames_to_column('genes') %>%
  filter(decision %in% c("Confirmed"))
# `TAGLN2`, `OIT3`, `IGFALS`, and `TTC36` common genes of rank 3 and rank 4 does not exist in 
#  confiremed cases of rank 6 
rank3 = rank3 %>%
  filter(!(genes %in% c('TAGLN2', 'OIT3', 'IGFALS','TTC36')))

rank4 = rank4 %>% 
  rownames_to_column('genes') %>%
  filter(decision %in% c("Confirmed"))
# `TAGLN2`, `OIT3`, `IGFALS`, and `TTC36` common genes of rank 3 and rank 4 does not exist in 
#  confiremed cases of rank 6 

rank4 = rank4 %>%
  filter(!(genes %in% c('TAGLN2', 'OIT3', 'IGFALS','TTC36')))



rank4[1:6,]
```

## Getting common features from ranked dataframes. 

#### Here, "common.ranking.3" has common genes with their ranking in HCCDB3 dataset and "common.ranking.4" has common genes with their ranking in HCCDB4 dataset. Common genes are ranked with different orders for HCCDB3 data and HCCDB4 data. Below codes checks these two dataframes iteratively and selects top 2, 4, 8, and all 16 genes considering both datasets.

```{r}

common = rank3[which(rank3$genes %in% rank4$genes),]



top=c()
top_list = list()

for (i in 1:dim(rank3)[1]){
  r3 = rank3[1:i,]
  r4 = rank4[1:i,]
  top = r4[which(r4$genes %in% r3$genes),"genes"]
  j = 1
  while ( 2^j <= dim(rank3)[1]){
    if (length(top) == 2^j) {
      name <- paste("top", 2^j, sep = "")
      assign(name, top)
      # cat("Top",2^j, "genes: ", top, '\n') # printing & checking each iteration
      top_list[[name]] <- top
    }
    j = j+1
  }
  # to add all common genes to the top list
  if (length(top) == dim(common)[1]){ 
      name <- paste("top", dim(common)[1], sep = "")
      assign(name, top)
      top_list[[name]] <- top
  }
}
top_list
write.csv( top_list$top8, "top_list1-8.csv", quote = F,row.names = F)

```

## Random Forest 
### Train: HCCDB3 Data (Train set)
### Test: HCCDB3 Data (Test set), HCCDB4 Data (Test set) 
### Validatation: HCCDB6 Data
##### Note: While using HCCDB4 as test set, only the part which is not used for Baruto Feature Selection is used. This condition is satisfied with set.seed () function for all data sets used.


#### Reading HCCDB3 data sets 
```{r}
x.3 = read.csv('DEGS_HCCDB3.csv', row.names = 1, header = TRUE)
x.3 = as.data.frame(t(x.3))

sample3_info = read.csv('prepared_sample_info3.csv', header = TRUE) 

head(sample3_info)
x.3[1:5,1:5]
```
### Scaling data (Data is already log2 transformed)
```{r}
x.3 = scale(x.3,center = TRUE, scale = TRUE)
x.3[1:5,1:5]
```
### Preparing data sets to split into test and train. Type varible (dependent variable) is added to the data as a factor.
```{r}
data3 = as.data.frame(x.3) %>% 
  rownames_to_column(var = "Samples") %>%
  left_join(sample3_info %>% select(Samples,TYPE), by = "Samples") %>%
  mutate(TYPE = as.factor(TYPE)) %>%
  column_to_rownames(var = "Samples")
```

### Generating a stratified random split of the data. 
```{r}
set.seed(11) # Always same train and test split will be used in the analysis
train3.Index <- createDataPartition(data3$TYPE, p = .7,
                                  list = FALSE,
                                  times = 1)
data3.train <- data3[train3.Index,]
test3 <- data3[-train3.Index,]
table(test3$TYPE)
table(data3.train$TYPE)

```

### Reading HCCDB4 data sets 
```{r}
x.4 = read.csv('DEGS_HCCDB4.csv', row.names = 1, header = TRUE)
x.4 = as.data.frame(t(x.4))

sample4_info = read.csv('prepared_sample_info4.csv', header = TRUE) 

head(sample4_info)
x.4[1:5,1:5]
```
### Scaling data (Data is already log2 transformed)
```{r}
x.4 = scale(x.4,center = TRUE, scale = TRUE)
x.4[1:5,1:5]
```
### Preparing data sets to split into test and train. Type varible is added to the data as a factor.
```{r}
data4 = as.data.frame(x.4) %>% 
  rownames_to_column(var = "Samples") %>%
  left_join(sample4_info %>% select(Samples,TYPE), by = "Samples") %>%
  mutate(TYPE = as.factor(TYPE)) %>%
  column_to_rownames(var = "Samples")
```

### Generating a stratified random split of the data
```{r}
set.seed(11) # Always same train and test split will be used in the analysis
train4.Index <- createDataPartition(data4$TYPE, p = .7,
                                  list = FALSE,
                                  times = 1)
data4.train <- data4[train4.Index,]
test4 <- data4[-train4.Index,]
table(test4$TYPE)
table(data4.train$TYPE)
```

### Reading HCCDB6 data sets 
```{r}
x.6 = read.csv('DEGS_HCCDB6.csv', row.names = 1, header = TRUE)
x.6 = as.data.frame(t(x.6))

sample6_info = read.csv('prepared_sample_info6.csv', header = TRUE) 

head(sample6_info)
x.6[1:5,1:5]
```

### Scaling data (Data is already log2 transformed)
```{r}
x.6 = scale(x.6,center = TRUE, scale = TRUE)
x.6[1:5,1:5]
```


### Preparing data sets to split into test and train. Type varible is added to the data as a factor.
```{r}
validation.6 = as.data.frame(x.6) %>% 
  rownames_to_column(var = "Samples") %>%
  left_join(sample6_info %>% select(Samples,TYPE), by = "Samples") %>%
  mutate(TYPE = as.factor(TYPE)) %>%
  column_to_rownames(var = "Samples")
```

```{r}
# A list that contains a number of custom named elements that the caret package looks for, such as how to fit and how to predict. 
# Define a custom RF (Classification) method to allow optimisation of ntree parameter as well as mtry.
customRF <- list(type = "Classification", library = "randomForest", loop = NULL)
customRF$parameters <- data.frame(parameter = c("mtry", "ntree"), 
                                  class = rep("numeric", 2), label = c("mtry", "ntree"))
customRF$grid <- function(x, y, len = NULL, search = "grid") {}
customRF$fit <- function(x, y, wts, param, lev, last, weights, ...) {
  randomForest(x, y, mtry = param$mtry, ntree=param$ntree, ...)
}
customRF$predict <- function(modelFit, newdata, preProc = NULL, submodels = NULL)
  predict(modelFit, newdata)
customRF$prob <- function(modelFit, newdata, preProc = NULL, submodels = NULL)
  predict(modelFit, newdata, type = "prob")
customRF$sort <- function(x) x[order(x[,1]),]
customRF$levels <- function(x) x$classes

# train model
control <- trainControl(method="repeatedcv", number=10, repeats=10, classProbs = TRUE)

```


```{r}

```


### Random Forest with top 2 genes 
```{r}

data3.train.top2 = data3.train %>%
  select(TYPE,top_list[["top2"]])

# Define values for mtry and ntree to explore. 
# For mtry assess 1 and 2 (since there are 2 genes)
tunegrid <- expand.grid(.mtry=c(1,2), 
                        .ntree=c(1000, 1500, 2000, 2500, 3000, 3500))
custom.top2 <- train(TYPE~., data=data3.train.top2, 
                 method=customRF, 
                 metric='Accuracy', 
                 tuneGrid=tunegrid, 
                 trControl=control, 
                 verbose=T) 
plot(custom.top2)

```


### Fitting the best model and Predicting on test 3, test 4 and validation set 6
```{r}
tunegrid <- expand.grid(.mtry = 1, .ntree =2500)
custom.top2 <- train(TYPE~., data=data3.train.top2, 
                 method=customRF, 
                 metric='Accuracy', 
                 tuneGrid=tunegrid, 
                 trControl=control, 
                 verbose=T) 
# Prepare Test sets and Validation set for top 2 genes
test3.top2 = test3 %>%
  select(TYPE,top_list[["top2"]])
test4.top2 = test4 %>%
  select(TYPE,top_list[["top2"]])
validation.6.top2 = validation.6 %>%
  select(TYPE,top_list[["top2"]])

prediction3 = predict(custom.top2,test3.top2,'prob')
prediction4 = predict(custom.top2,test4.top2,'prob')
prediction6 = predict(custom.top2,validation.6.top2,'prob')
```

```{r}
pdf(file = "Combination 1 - ROC Curves for 2 Genes.pdf",   # The directory you want to save the file in
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

par(pty="s") 
lrROC <- roc(test3.top2$TYPE ~ prediction3$HCC,plot=TRUE,
             print.auc=T,print.auc.x = 0.9,print.auc.y = 0.75,
             col="red",
             lwd =1.5,legacy.axes=TRUE,main="ROC Curves for 2 Genes")
lrROC <- roc(test4.top2$TYPE ~ prediction4$HCC,add = TRUE,plot=TRUE,
             print.auc=T,print.auc.x = 0.4,print.auc.y = 0.93,
             col="orange",
             lwd =1.5,legacy.axes=TRUE)
lrROC <- roc(validation.6.top2$TYPE ~ prediction6$HCC,add = TRUE,plot=TRUE,
             print.auc=T,print.auc.x = 0.9,print.auc.y = 0.45,
             col="blue",
             lwd =1.5,legacy.axes=TRUE)
legend("bottomright",cex = 0.65,xjust = 1,legend=c('HCCDB3 Test set','HCCDB4 Test set','HCCDB6 Validation Set',
                                                   "mtry = 1","ntree = 2500"),
       col=c('red','orange','blue','black','black'),lwd=1)
dev.off()

```


### Random Forest with top 4 genes 
```{r}
data3.train.top4 = data3.train %>%
  select(TYPE,top_list[["top4"]])
mtry <- sqrt(ncol(data3.train.top4)-1)
# Define values for mtry and ntree to explore. 
# For mtry assess 50%, 100%, and 150% of the default value.
tunegrid <- expand.grid(.mtry=c(round(0.5*mtry),round(mtry),round(1.5*mtry)), 
                        .ntree=c(1000, 1500, 2000, 2500, 3000))
custom.top4 <- train(TYPE~., data=data3.train.top4, 
                 method=customRF, 
                 metric='Accuracy', 
                 tuneGrid=tunegrid, 
                 trControl=control, 
                 verbose=T) 
plot(custom.top4)
```

### Fitting the best model and Predicting on test 3, test 4 and validation set 6
```{r}
tunegrid <- expand.grid(.mtry = round(0.5*mtry), .ntree =1000)
custom.top4 <- train(TYPE~., data=data3.train.top4, 
                 method=customRF, 
                 metric='Accuracy', 
                 tuneGrid=tunegrid, 
                 trControl=control, 
                 verbose=T) 
# Prepare Test sets and Validation set for top 4 genes
test3.top4 = test3 %>%
  select(TYPE,top_list[["top4"]])
test4.top4 = test4 %>%
  select(TYPE,top_list[["top4"]])
validation.6.top4 = validation.6 %>%
  select(TYPE,top_list[["top4"]])
prediction3 = predict(custom.top4,test3.top4,'prob')
prediction4 = predict(custom.top4,test4.top4,'prob')
prediction6 = predict(custom.top4,validation.6.top4,'prob')

```

```{r}
pdf(file = "Combination 1 - ROC Curves for 4 Genes.pdf",   # The directory you want to save the file in
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

par(pty="s") 
lrROC <- roc(test3.top4$TYPE ~ prediction3$HCC,plot=TRUE,
             print.auc=T,print.auc.x = 0.9,print.auc.y = 0.75,
             col="red",
             lwd =1.5,legacy.axes=TRUE,main="ROC Curves for 4 Genes")
lrROC <- roc(test4.top4$TYPE ~ prediction4$HCC,add = TRUE,plot=TRUE,
             print.auc=T,print.auc.x = 0.4,print.auc.y = 0.93,
             col="orange",
             lwd =1.5,legacy.axes=TRUE)
lrROC <- roc(validation.6.top4$TYPE ~ prediction6$HCC,add = TRUE,plot=TRUE,
             print.auc=T,print.auc.x = 0.9,print.auc.y = 0.45,
             col="blue",
             lwd =1.5,legacy.axes=TRUE)
legend("bottomright",cex = 0.65,xjust = 1,legend=c('HCCDB3 Test set','HCCDB4 Test set','HCCDB6 Validation Set',
                                                   "mtry = 1","ntree = 1000"),
       col=c('red','orange','blue','black','black'),lwd=1)
dev.off()

```

### Random Forest with top 8 genes 
```{r}
data3.train.top8 = data3.train %>%
  select(TYPE,top_list[["top8"]])
mtry <- sqrt(ncol(data3.train.top8)-1)
# Define values for mtry and ntree to explore. 
# For mtry assess 50%, 100%, and 150% of the default value.
tunegrid <- expand.grid(.mtry=c(round(0.5*mtry),round(mtry),round(1.5*mtry)), 
                        .ntree=c(1000, 1500, 2000, 2500, 3000))
custom.top8 <- train(TYPE~., data=data3.train.top8, 
                 method=customRF, 
                 metric='Accuracy', 
                 tuneGrid=tunegrid, 
                 trControl=control, 
                 verbose=T) 
plot(custom.top8)
```

### Fitting the best model and Predicting on test 3, test 4 and validation set 6
```{r}
tunegrid <- expand.grid(.mtry = round(0.5*mtry), .ntree =3000)
custom.top8 <- train(TYPE~., data=data3.train.top8, 
                 method=customRF, 
                 metric='Accuracy', 
                 tuneGrid=tunegrid, 
                 trControl=control, 
                 verbose=T) 
# Prepare Test sets and Validation set for top 8 genes
test3.top8 = test3 %>%
  select(TYPE,top_list[["top8"]])
test4.top8 = test4 %>%
  select(TYPE,top_list[["top8"]])
validation.6.top8 = validation.6 %>%
  select(TYPE,top_list[["top8"]])
prediction3 = predict(custom.top8,test3.top8,'prob')
prediction4 = predict(custom.top8,test4.top8,'prob')
prediction6 = predict(custom.top8,validation.6.top8,'prob')

```

```{r}
pdf(file = "Combination 1 - ROC Curves for 8 Genes.pdf",   # The directory you want to save the file in
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

par(pty="s") 
lrROC <- roc(test3.top8$TYPE ~ prediction3$HCC,plot=TRUE,
             print.auc=T,print.auc.x = 0.9,print.auc.y = 0.75,
             col="red",
             lwd =1.5,legacy.axes=TRUE,main="ROC Curves for 8 Genes")
lrROC <- roc(test4.top8$TYPE ~ prediction4$HCC,add = TRUE,plot=TRUE,
             print.auc=T,print.auc.x = 0.4,print.auc.y = 0.93,
             col="orange",
             lwd =1.5,legacy.axes=TRUE)
lrROC <- roc(validation.6.top8$TYPE ~ prediction6$HCC,add = TRUE,plot=TRUE,
             print.auc=T,print.auc.x = 0.9,print.auc.y = 0.45,
             col="blue",
             lwd =1.5,legacy.axes=TRUE)
legend("bottomright",cex = 0.65,xjust = 1,legend=c('HCCDB3 Test set','HCCDB4 Test set','HCCDB6 Validation Set',
                                                   "mtry = 1","ntree = 3000"),
       col=c('red','orange','blue','black','black'),lwd=1)
dev.off()
```

### Random Forest with top 16 genes 
```{r}
data3.train.top16 = data3.train %>%
  select(TYPE,top_list[["top16"]])
mtry <- sqrt(ncol(data3.train.top16)-1)
# Define values for mtry and ntree to explore. 
# For mtry assess 50%, 100%, and 150% of the default value.
tunegrid <- expand.grid(.mtry=c(round(0.5*mtry),round(mtry),round(1.5*mtry)), 
                        .ntree=c(1000, 1500, 2000, 2500, 3000))
custom.top16 <- train(TYPE~., data=data3.train.top16, 
                 method=customRF, 
                 metric='Accuracy', 
                 tuneGrid=tunegrid, 
                 trControl=control, 
                 verbose=T) 
plot(custom.top16)
```

### Fitting the best model and Predicting on test 3, test 4 and validation set 6
```{r}
tunegrid <- expand.grid(.mtry = round(1.5*mtry), .ntree =1500)
custom.to16 <- train(TYPE~., data=data3.train.top16, 
                 method=customRF, 
                 metric='Accuracy', 
                 tuneGrid=tunegrid, 
                 trControl=control, 
                 verbose=T) 
# Prepare Test sets and Validation set for top 16 genes
test3.top16 = test3 %>%
  select(TYPE,top_list[["top16"]])
test4.top16 = test4 %>%
  select(TYPE,top_list[["top16"]])
validation.6.top16 = validation.6 %>%
  select(TYPE,top_list[["top16"]])
prediction3 = predict(custom.top16,test3.top16,'prob')
prediction4 = predict(custom.top16,test4.top16,'prob')
prediction6 = predict(custom.top16,validation.6.top16,'prob')

```

```{r}
pdf(file = "Combination 1 - ROC Curves for 16 Genes.pdf",   # The directory you want to save the file in
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

par(pty="s") 
lrROC <- roc(test3.top16$TYPE ~ prediction3$HCC,plot=TRUE,
             print.auc=T,print.auc.x = 0.9,print.auc.y = 0.75,
             col="red",
             lwd =1.5,legacy.axes=TRUE,main="ROC Curves for 16 Genes")
lrROC <- roc(test4.top16$TYPE ~ prediction4$HCC,add = TRUE,plot=TRUE,
             print.auc=T,print.auc.x = 0.4,print.auc.y = 0.93,
             col="orange",
             lwd =1.5,legacy.axes=TRUE)
lrROC <- roc(validation.6.top16$TYPE ~ prediction6$HCC,add = TRUE,plot=TRUE,
             print.auc=T,print.auc.x = 0.9,print.auc.y = 0.45,
             col="blue",
             lwd =1.5,legacy.axes=TRUE)
legend("bottomright",cex = 0.65,xjust = 1,legend=c('HCCDB3 Test set','HCCDB4 Test set','HCCDB6 Validation Set',
                                                   "mtry = 6","ntree = 1500"),
       col=c('red','orange','blue','black','black'),lwd=1)
dev.off()
```



### Random Forest with top 23 genes 
```{r}
data3.train.top23 = data3.train %>%
  select(TYPE,top_list[["top23"]])
mtry <- sqrt(ncol(data3.train.top23)-1)
# Define values for mtry and ntree to explore. 
# For mtry assess 50%, 100%, and 150% of the default value.
tunegrid <- expand.grid(.mtry=c(round(0.5*mtry),round(mtry),round(1.5*mtry)), 
                        .ntree=c(1000, 1500, 2000, 2500, 3000))
custom.top23 <- train(TYPE~., data=data3.train.top23, 
                 method=customRF, 
                 metric='Accuracy', 
                 tuneGrid=tunegrid, 
                 trControl=control, 
                 verbose=T) 
plot(custom.top23)
```



### Fitting the best model and Predicting on test 3, test 4 and validation set 6
```{r}
tunegrid <- expand.grid(.mtry = round(mtry), .ntree =1500)
custom.top23 <- train(TYPE~., data=data3.train.top23, 
                 method=customRF, 
                 metric='Accuracy', 
                 tuneGrid=tunegrid, 
                 trControl=control, 
                 verbose=T) 
# Prepare Test sets and Validation set for top 16 genes
test3.top23 = test3 %>%
  select(TYPE,top_list[["top23"]])
test4.top23 = test4 %>%
  select(TYPE,top_list[["top23"]])
validation.6.top23 = validation.6 %>%
  select(TYPE,top_list[["top23"]])
prediction3 = predict(custom.top23,test3.top23,'prob')
prediction4 = predict(custom.top23,test4.top23,'prob')
prediction6 = predict(custom.top23,validation.6.top23,'prob')

```

```{r}
pdf(file = "Combination 1 - ROC Curves for 23 Genes.pdf",   # The directory you want to save the file in
    width = 11, # The width of the plot in inches
    height = 8.5) # The height of the plot in inches

par(pty="s") 
lrROC <- roc(test3.top23$TYPE ~ prediction3$HCC,plot=TRUE,
             print.auc=T,print.auc.x = 0.9,print.auc.y = 0.75,
             col="red",
             lwd =1.5,legacy.axes=TRUE,main="ROC Curves for 23 Genes")
lrROC <- roc(test4.top23$TYPE ~ prediction4$HCC,add = TRUE,plot=TRUE,
             print.auc=T,print.auc.x = 0.4,print.auc.y = 0.93,
             col="orange",
             lwd =1.5,legacy.axes=TRUE)
lrROC <- roc(validation.6.top23$TYPE ~ prediction6$HCC,add = TRUE,plot=TRUE,
             print.auc=T,print.auc.x = 0.9,print.auc.y = 0.45,
             col="blue",
             lwd =1.5,legacy.axes=TRUE)
legend("bottomright",cex = 0.65,xjust = 1,legend=c('HCCDB3 Test set','HCCDB4 Test set','HCCDB6 Validation Set',
                                                   "mtry = 5","ntree = 1500"),
       col=c('red','orange','blue','black','black'),lwd=1)
dev.off()
```


